Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-09-08 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:09:08	22:34:55	(Tue 8.Sep)

Adding slightly steered versions of each window to the window database gives LCA greater flexibility in searching for optimal window. A Kaiser window $\f_\beta$ steered to the angle $\phi$ can be expressed as 
%
\begin{align}
\w_{\beta,\phi} = \frac{\f_\beta\H\diag{(\a_\phi)}}{\f_\beta\H\a_\phi}
\end{align}
%
where diag($\a_\phi$) is a diagonal matrix constructed from the steering vector $\a_\phi$:
%
\begin{align}
\a_\phi = \bmat{
1 \\
e^{-j\frac{md}{\lambda}\sin(\phi)} \\
\vdots\\
e^{-j\frac{(M-1)d}{\lambda}\sin(\phi)}
}.
\end{align}
%
Here $d$ is the element spacing, $\lambda$ is the wavelength and $\phi$ is the steering amount. The normalization factor $\f_\beta\H\a_\phi$ is the reciprocal of the window's coherent gain. It ensures unit gain in the direction of interest as required by (\ref{eq_lca_definition}). Since the signal-to-noise ratio is constant for a specific window (same $\beta$-value), this normalization also proportionally increases the incoherent noise gain.

@@

\section{Results and Discussion}

@@

\begin{itemize}
\item More steering -> more aggressive suppression of interference at the cost of boosted sidelobes / noise.
\item More steering -> Larger variations in amplitude of sources
\end{itemize}

@@

So far we have determined it reasonable to use Kaiser windows in the range $\beta\in[0,5]$, steered within the range $|\phi|\in[0,\phi_\mathrm{3dB}]$. But many windows do we need in these ranges?

@@

Observe that sampling the $\beta$-range more finely increases the sidelobe suppression but leaves resolution unchanged. Adding more steering-variations improves both sidelobe suppression and resolution.

However, using more than  $N_\beta=2$ window types and $N_\beta=3$ steering angles makes minimal

\ref{oversampling_mosaic}

From our observations one should have at least twice the number of beams as channels. Image is more visually pleasing and appears more detailed up to $8M$.

@@

TODO

@@

>>>>

Adding slightly steered versions of each window to the window database gives LCA greater flexibility in searching for optimal window. A Kaiser window $\f_\beta$ steered to the angle $\phi$ can be expressed as 
%
\begin{align}
\w_{\beta,\phi} = \frac{\f_\beta\H\diag{(\a_\phi)}}{\f_\beta\H\a_\phi}
\end{align}
%
where diag($\a_\phi$) is a diagonal matrix constructed from the steering vector $\a_\phi$:
%
\begin{align}
\a_\phi = \bmat{
1 \\
e^{-j\frac{2\pi d}{\lambda}\sin(\phi)} \\
\vdots\\
e^{-j\frac{2\pi (M-1)d}{\lambda}\sin(\phi)}
}.\label{steering_vector}
\end{align}
%
Here $d$ is the element spacing, $\lambda$ is the wavelength and $\phi$ is the steering amount. The normalization factor $\f_\beta\H\a_\phi$ is the reciprocal of the window's coherent gain. It ensures unit gain in the direction of interest as required by (\ref{eq_lca_definition}). Since the signal-to-noise ratio is constant for a specific window (same $\beta$-value), this normalization also proportionally increases the incoherent noise gain.

@@

\section{Results and Discussion}\label{results_discussion}

@@

@@

So far we have determined it reasonable to use Kaiser windows in the range $\beta\in[0,5]$, steered within the range $|\phi|\in[0,\phi_\mathrm{3dB}]$. But how many windows do we need in these ranges?

@@

Observe that sampling the $\beta$-range more finely increases the sidelobe suppression but leaves resolution unchanged. Adding more steering-variations improves both sidelobe suppression and resolution. We can observe a major improvement going to $N_\beta=2$ and $N_\phi=3$, but minimal improvement by adding yet more windows.

@@

The computational complexity of MVDR is generally considered to be of O($M^3$). However, the implementation we are using is optimized so that building the spatial covariance matrix is of O($N_kN_L,L^2$), and inverting it is of O($L^3$). We describe this implementation in \cite{Buskenes2014}, and a beamspace version in \cite{Asen2013}, where we optimize if for a graphics processing unit (GPU) for nearly two orders of magnitude speed increase compared to a straightforward C implementation. However, MVDR is not ideal for GPUs due to the complex data dependencies present in the covariance computation, and at best we only managed to utilize 10\% of the GPUs potential.

In comparison, LCA is of O($M N_w$), and extremely well suited for GPUs. Each pixel depends on only a small subset of data, and the windows can be precomputed and stored in GPU cache for near immediate access. Furthermore, since we have observed LCA to perform similarly using other parametric window variants, the complexity can be reduced to near DAS.

@@

\section{Reducing LCA complexity}

\renewcommand\b{\vec b}

Now let us choose a trigonometric window function:
%
\begin{align}
\w = e^{j\phi_\theta[m]}\big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\big)
\end{align}
%
where $\phi_\theta[m] = \frac{2\pi}{\lambda}dm\sin\theta$. Inserting this yields:
%
\begin{align}
\w\H\x &= \sumb{m=0}{M-1} e^{j\phi_\theta[m]}\Big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\Big) x_m^* \nn
&= \alpha\sumb{m=0}{M-1} e^{j\phi_\theta[m]} x_m^* - (1-\alpha)\sumb{m=0}{M-1} e^{j\varphi[m]}\cos\left(\frac{2\pi m}{M-1}\right) x_m^* \nn
&= \alpha \a\T\x + (1-\alpha)\b\T\x
\end{align}
%
where $\a_\phi$ is the steering vector defined in (\ref{steering_vector}), and 
%
\begin{align}
\a = \begin{bmatrix}
     1 \\
     e^{j\frac{2\pi}{\lambda}d\sin\theta} \\
     e^{j\frac{2\pi}{\lambda}2d\sin\theta} \\
     \vdots \\
     e^{j\frac{2\pi}{\lambda}d(M-1)\sin\theta}
     \end{bmatrix} \qquad \text{and} \qquad
\b = \begin{bmatrix}
     0 \\
     e^{j\frac{2\pi}{\lambda}d\sin\theta}\sin\left(\frac{2\pi m}{M-1}\right) \\
     e^{j\frac{2\pi}{\lambda}2d\sin\theta}\sin\left(\frac{2\pi m}{M-1}\right) \\
     \vdots \\
     0
     \end{bmatrix}\a
\end{align}
%
To attain unity gain in the look direction the weights must sum to one. 
%
\begin{align}
\w\T\1 &= \big(\alpha\a - (1-\alpha)\b\big)\T\1 = 1
\end{align}
%
This is true for any value of $\alpha$ if $\a\T\1 = 1$ and $\b\T\1 = 1$.

where
%
\begin{align}
|\w\H\x|^2 &= \Big| \alpha a + (1-\alpha)b \Big|^2 \nn
&= \Big[\alpha a + (1-\alpha)b\Big] \Big[\alpha a^* + (1-\alpha)b^*\Big] \nn
&= \alpha^2 aa^* + \alpha a(1-\alpha)b^* + \alpha a^*(1-\alpha)b + (1-\alpha)^2bb^* \nn
&= \alpha^2 aa^* + \alpha(1-\alpha)(ab^* + a^*b) + (1-\alpha)^2bb^* \nn
&= \alpha^2 (aa^* - (ab^* + a^*b) - bb^*) + \alpha(ab^* + a^*b) + bb^* \nn
\end{align}
%
\begin{align}
\frac{\partial}{\partial\alpha} |\w\H\x|^2 
&= 2\alpha (aa^* - (ab^* + a^*b) - bb^*) + ab^* + a^*b = 0
\end{align}
%
\begin{align}
\alpha &= \frac{ab^* + a^*b}{2(aa^* - (ab^* + a^*b) - bb^*)}
\end{align}

LCA can then be defined as
%
\begin{align}
\argmin{\w}\ |z|^2 = \argmin{\w}\ z\,z^* = %\argmin{\w}\ \w\H\x = \sumb{m=0}{M-1} w_m^* x_m
\end{align}
